---
title: "Zementis Introduction - Training Exercises"
tutorial:
  version: 1.1
author: Alexander Lemm
output: 
  learnr::tutorial:
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(httr)
library(jsonlite)
library(dplyr)

tutorial_options(exercise.completion = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```

## Model Deployment

### Exercises

#### Exercise 1

Navigate to the Zementis demo server at https://myadapa.zementis.com/adapaconsole/console.jsf and log in using your credentials you received earlier via email.

```{r deploy-one-quiz}

question("How many prediction models were already uploaded to your instance and are available to you?",
    answer("8"),
    answer("3", correct = TRUE),
    answer("1")
  )
```

#### Exercise 2

Now, it is time to upload a new prediction model: You will upload the credit risk model you created in the Demystifying Data Science course. 

In your Zementis instance click on `Upload PMML Files` and select the credit risk model `credit_risk_classification_tree.xml` on your disk which resides in `~/DDS_course_export`.


```{r deploy-two-quiz}

question("What message is shown in the web console after uploading your model?",
    answer("The provided file was uploaded successfully.", correct = TRUE),
    answer("The provided PMML file was uploaded successfully.")
  )
```

## Predicting New Values

### Exercises

#### Exercise 1

Download `loan_data_new.csv` which is part of the `zementisTutorials` package by clicking on the following link: `r downloadLink("downloadCSV", "Get CSV file")`


```{r download-csv, context="server"}

file_name <- "loan_data_new.csv"

output$downloadCSV <- downloadHandler(
  filename = file_name,
  content = function(file) {
    file.copy(file.path( system.file("tutorials/zementis_introduction/data", package = "zementisTutorials"), 
      file_name), 
      file)
  }
)
```

In your Zementis server instance navigate to `Score/Classify Data`. Select your credit risk model from the model list. Next, upload `loan_data_new.csv` to the Zementis server.


```{r predict-upload-one-quiz}

question("According to the predictions would you accept the loan request of the first applicant?",
    answer("The model scored 0 for the first applicant. So we would accept the request."),
    answer("The model scored 1 for the first applicant, so we would deny the request.", correct = TRUE)
  )
```

#### Exercise 2

**Note**: To answer this question properly you might need to modify the settings so that all prediction results are shown on screen and not just the first 10.

```{r predict-upload-two-quiz}

question("According to the predictions would you accept the loan request of the last applicant?",
    answer("The model scored 0 for the last applicant. So we would accept the request.", correct = TRUE),
    answer("The model scored 1 for the last applicant, so we would deny the request.")
  )
```



#### Exercise 3

In this exercise you will use the embedded REST API client on the server which is part of the API documentation to test a single GET request

* In your browser, please navigate to the REST API page at https://myadapa.zementis.com/adapars/ and enter your credentials
* On the page, please open the `GET /apply/{model_name}` section which describes how to score a single instance via the REST API
* In the `model name` window please enter `Credit Risk Model`
* Copy the JSON code you see below to the `record` window:

```
{
    "loan_amnt": 5000,
    "grade": "G",
    "home_ownership": "MORTGAGE",
    "annual_inc": 250000,
    "age": 30,
    "emp_bin": "[0,1]",
    "int_bin": "(7.51,10.2]"
}

```

* Click on `Try it out`

```{r predict-rest-one-quiz}

question("Check out the prediction returned from the server. Would you accept this loan request?",
    answer("The model predicted that the loan will be paid back. We would accept the request."),
    answer("The model predicted that the loan won't be paid back, We would not accept the request.", correct = TRUE)
  )
```



#### Exercise 4

Now, you will use an external REST API client to send a GET request to your EC2 instance. Luckily, we can use the R console below for this. The code in the console below represents a very lean client and will send a loan request of a single applicant to your server using GET. 

Before clicking on `Run Code` in the console, please add your Zementis server credentials to the GET request otherwise you won't receive any results.


```{r predict-rest-client, exercise=TRUE}

body <- '{
    "loan_amnt": 5000,
    "grade": "A",
    "home_ownership": "MORTGAGE",
    "annual_inc": 250000,
    "age": 30,
    "emp_bin": "[0,1]",
    "int_bin": "(7.51,10.2]"
}'

GET("https://myadapa.zementis.com:443/adapars/apply/Credit%20Risk%20Model",
    authenticate("YOUR_USERNAME", "YOUR_PASSWORD"),
    query = list(record = body),
    content_type_json()) %>% content(as = "parsed") %>% toJSON(pretty = TRUE) 
```


```{r predict-rest-client-quiz}

question("Check out the prediction returned from the server. Would you accept this loan request?",
    answer("The model predicted that the loan will be paid back. We would accept the request.", correct = TRUE),
    answer("The model predicted that the loan won't be paid back, We would not accept the request.")
  )
```


#### Exercise 5

Go ahead and install the Zementis Excel Add-In on your machine. You should have received the respective instructions and the installation file earlier via email.

#### Exercise 6

In this final exercise you will send the entire loan request data included in `plugin_data.xlsx` which is part of the `zementisTutorials` package.

* Download `plugin_data.xlsx` by clicking on the following link: `r downloadLink("downloadExcel", "Get Excel")`
* Open `plugin_data.xlsx` on your machine using Excel
* Navigate to the `Predict with ADAPA` tab
* Please enter the respective data under `Setup connection`(Go with the default URL by clicking on `Use demo`)
* Select the entire data of the sheet
* Choose the credit risk model in the `Apply Model` section and send the data to the Zementis server

```{r download-excel, context="server"}

excel_name <- "plugin_data.xlsx"

output$downloadExcel <- downloadHandler(
  filename = excel_name,
  content = function(file) {
    file.copy(file.path( system.file("tutorials/zementis_introduction/data", package = "zementisTutorials"), 
      excel_name), 
      file)
  }
)
```


```{r predict-excel-quiz}
quiz(
  question("Would you accept the loan request of the 12th applicant?",
           answer("The model predicted that the loan will be paid back. We would accept
                  the request."),
           answer("The model predicted that the loan won't be paid back, We would no
                  accept the request.", correct = TRUE)
  ),
    question("Would you accept the loan request of the 6th applicant?",
           answer("The model predicted that the loan will be paid back. We would accept
                  the request.", correct = TRUE),
           answer("The model predicted that the loan won't be paid back, We would no
                  accept the request.")
  )
  
)
```

